---
title: "Effects of roads and forest successional age on species in an experimental plant invasion"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: yes
  word_document: null
---

```{r set options, include=F, echo=F}

knitr::opts_chunk$set(fig.width = 7.5)
knitr::opts_chunk$set(dpi = 600)
knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(message = F)
# knitr::opts_chunk$set(eval = F)

# options(na.action = "na.fail")# global model setting for using multimodel inference 'dredge' function

```

```{r load packages}
# library(lme4); library(car); library(sciplot); library(lmerTest); 
# library(MuMIn)
library(plyr); library(dplyr); library(tidyr); library(readr)
library(ggplot2)
library(knitr)
```

```{r load data}
seedling_data <- read_csv("data/seedling_growth_survival_data.csv")
germination_data <- read_csv("data/germination_data.csv")
forest_data <- read_csv("data/forest_trees.csv")
light_data <- read_csv("data/light2005_data.csv")
litter_data <- read_csv("data/litter.csv")
soil_moisture_data <- read_csv("data/soil_moisture2005_data.csv")
```

# 

```{r check data}
# summary(seedling_data)
# unique(seedling_data$species)
seedling_data <- filter(seedling_data, is.na(species)=="FALSE")
# filter(seedling_data, rgr_ht<0)
seedling_data <- seedling_data %>% 
  mutate(rgr_ht = (final_ht06 - initial_ht)/initial_ht)
## There are 97 negative values in this growth rate. Values are sometimes very inconsistent from the spring 2006 measurement to the final measurement

# summary(germination_data)
# unique(germination_data$species)
# filter(germination_data, is.na(seeds_germinated))$comments
germination_data <- germination_data %>% 
  mutate(pct_germinated = 100*(seeds_germinated/seeds_present),
         common_name = "",
         status = native_exotic)

# summary(forest_data)
# unique(forest_data$species)
# filter(forest_data, is.na(species))

# summary(light_data)

# summary(litter_data)

# summary(soil_moisture_data)

# xtabs(~ species + native_exotic, germination_data)
```

```{r make species table}
germination_data$common_name[germination_data$species=="D"] <- "Dogwood"
germination_data$common_name[germination_data$species=="A"] <- "Autumn olive"
germination_data$common_name[germination_data$species=="H"] <- "Honeysuckle"
germination_data$common_name[germination_data$species=="P"] <- "Privet"
germination_data$common_name[germination_data$species=="R"] <- "Redbud"
germination_data$common_name[germination_data$species=="S"] <- "Spicebush"

germination_data$status[germination_data$status=="N"] <- "Native"
germination_data$status[germination_data$status=="E"] <- "Non-native"

species_table <- germination_data %>% 
  filter(site=="BB", distance==0) %>%
  select(species, status, common_name)
  
kable(species_table %>% 
        select(Species = common_name, Status = status) %>%
        arrange(Status)
      )
```


## Experiment A: Seed germination

* Ten seeds from each of six species (3 invaders, 3 natives) were placed at each distance


## Experiment B: Seedling survival and growth

* Seedlings of six species (3 invaders, 3 natives) were planted at 12 forest sites classified as mature (>60 years) or young (<40 years).  

* Four seedlings of each species at distances of 0, 5, 10, 20, 40, and 60 meters from the roadway edge. 

* All seedlings were 10-15cm tall when planted

**H1:** There is a negative effect on survival and growth for all species moving from edge to interior

* Performance will also be poorer in mature versus young forests
    

**H2:** Non-native species perform relatively better near the edge and in young forests.

* Native species perform relatively better in interior locations and in mature forests


